---
title: "ISCI448"
author: "Olivia T. Li and Katrina M. Kaur"
date: "April 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Species interactions (for example, between plants and their pollinators or predators and their prey) play a major role in shaping both a species’ ecology (their abundance and distribution) and their evolutionary trajectories. 

<<<<<<< HEAD
In this project, we focus specifically on how these interactions drive evolution over very long time periods (e.g., millions of years). While there is a great deal of theory and empirical evidence supporting the importance of competitive interactions between species on shaping large-scale patterns of evolution, there is relatively little known about how the consequences of engaging in mutualistic/cooperative interactions. Mathematical theory suggests that cooperating with another species should reduce the chance that one of the partners ‘speciates’ (splits into two new species); as a result non-mutualistic species should become more common than the opposite. 
=======
In this project, we focus specifically on how these interactions drive evolution over very long time periods (e.g., millions of years). While there is a great deal of theory and empirical evidence supporting the importance of competitive interactions between species on shaping large-scale patterns of evolution, there is relatively little known about how the consequences of engaging in mutualistic/cooperative interactions. Mathematical theory suggests that cooperating with another species should reduce the chance that one of the partners ‘speciates’ (splits into two new species); as a result non-mutualistic species should become more common than the opposite. However, a number of recent empirical studies have found that groups that engage in cooperation are actually more diverse than groups that do not; this observation is at odds with the theory. In this project we have tried to reconcile these contradiction by investigating in finer detail the structure of the network of our study group (ants and plants). In order to do this, we have constructed a network showing which ant species interact with which plant species.
>>>>>>> 45d793efa0234913fab763262c849e71ea9c3c9b

However, a number of recent empirical studies have found that groups that engage in cooperation are actually more diverse than groups that do not; this observation is at odds with the theory. In this project we have tried to reconcile these contradiction by investigating in finer detail the structure of the network of our study group (ants and plants). In order to do this, we have constructed a network showing which ant species interact with which plant species.

We have used computer code to “read” through thousands and thousands of scientific abstracts and pull out information on which ant species are involved in mutualistic interactions with plants (as well as which species of plants). We have then assembled this information into a database so that a network of plant-ant interactions can be built. This network will then be used in the future to address questions regarding whether mutualistic/cooperative interactions generate diversity.

## Packages used
```{r}
library(tidyr)
library(dplyr)
library(zoo)
library(taRifx)
```

## Step 1: Data Organization 
First we need to convert the raw data into a more workable format. To do so, we have grouped the ant species and plant species together according to their abstract number for each of the three characteristics. 

## SEED DISPERSAL 

## Separating columns for abstract number and genus and species name
## Leaving NA on the left side if the abstract number is missing
```{r}
sd_raw <- read.csv(file = "sd_network_tm_raw_2.csv", header = TRUE)

sd_tidy_table <- 
      sd_raw %>%
        separate(abs.name, into = c("abs", "genus", "species"), fill = "left")
```

## Creating new column with full binomial name, then removing the genus and species columns
```{r}
sd_tidy_table$name <- paste(sd_tidy_table$genus, sd_tidy_table$species)
sd_tidy_table$genus <- NULL
sd_tidy_table$species <- NULL
```

## Filling in the NA with the abstracts above
```{r}
sd_tidy_table2 <- sd_tidy_table %>% 
                  do(na.locf(.))
write.csv(sd_tidy_table2, file = "sd_network_tm_processed_2.csv")
```

## Data is now prepared for process_data.R file to make 0 and 1 matrix
## Repeat the same processes for the two other characteristics below


## EXTRAFLORAL NECTAR
```{r}
efn_raw <- read.csv(file = "efn_network_tm_raw_2.csv", header = TRUE)

efn_tidy_table <- 
  efn_raw %>%
  separate(abs.name, into = c("abs", "genus", "species"), fill = "left")

efn_tidy_table$name <- paste(efn_tidy_table$genus, efn_tidy_table$species)
efn_tidy_table$genus <- NULL
efn_tidy_table$species <- NULL

efn_tidy_table2 <- efn_tidy_table %>% 
  do(na.locf(.))
write.csv(efn_tidy_table2, file = "efn_network_tm_processed_2.csv")
```


#DOMATIA
```{r}
dom_raw <- read.csv(file = "dom_network_tm_raw_2.csv", header = TRUE)

dom_tidy_table <- 
  dom_raw %>%
  separate(abs.name, into = c("abs", "genus", "species"), fill = "left")

dom_tidy_table$name <- paste(dom_tidy_table$genus, dom_tidy_table$species)
dom_tidy_table$genus <- NULL
dom_tidy_table$species <- NULL

dom_tidy_table2 <- dom_tidy_table %>% 
  do(na.locf(.))
write.csv(dom_tidy_table2, file = "dom_network_tm_processed_2.csv")
```



## Process_data blurb

## Loading list of ants and plants to use in sorting the sd, efn, dom lists
## Ants
```{r}
ant_list <- read.csv("worldants_specieslist.csv", header=TRUE)
ant_list <- as.data.frame(ant_list$current.valid.name)
colnames(ant_list) <- "ants"
levels(ant_list$ants) <- tolower(levels(ant_list$ants))
```

## Plants
```{r}
angiosperms_list <- as.data.frame(read.csv("all_angiosperms.csv", header = TRUE, stringsAsFactors = FALSE))
colnames(angiosperms_list) <- "plants"
```


## SEED DISPERSAL

## Removing abstracts that only have one species appear
```{r}
sd_network <- read.csv("sd_network_tm_processed_2.csv", header=TRUE)
sd_network2 <- sd_network[sd_network$abs %in% sd_network$abs[duplicated(sd_network$abs)],]
sd_network2$X <- NULL
```

## Sorting into two lists, one with the ants that appeared and one with the plants that appeared
```{r}
ants_sd <- sd_network2 %>% filter(name %in% ant_list$ants)
ants_sd <- remove.factors(ants_sd)
plants_sd <- sd_network2 %>% filter(name %in% angiosperms_list$plants)
plants_sd <- remove.factors(plants_sd)

write.csv(ants_sd, file = "seed_dispersal_ants_2.csv")
write.csv(plants_sd, file = "seed_dispersal_plants_2.csv")
```

## Merging into one table with the columns labelled abstracts, ants, and plants
```{r}
together_sd <- merge(ants_sd, plants_sd, by="abs")
colnames(together_sd) <- c("abs", "ants", "plants")
```

## Removing abstract IDs as they are no longer needed
```{r}
together_sd$abs <- NULL
```

## Turning table with two columns into a ) and 1 matrix
```{r}
final_table_sd <- table(together_sd)
final_table_sd <- (final_table_sd > 0) + 0
write.csv(final_table_sd, file = "sd_01_network.csv")
```

## Repeat the same processes for the two other characteristics below


## EXTRAFLORAL NECTAR
```{r}
efn_network <- read.csv("efn_network_tm_processed_2.csv", header=TRUE)
efn_network2 <- efn_network[efn_network$abs %in% efn_network$abs[duplicated(efn_network$abs)],]
efn_network2$X <- NULL

ants_efn <- efn_network2 %>% filter(name %in% ant_list$ants)
ants_efn <- remove.factors(ants_efn)
plants_efn <- efn_network2 %>% filter(name %in% angiosperms_list$plants)
plants_efn <- remove.factors(plants_efn)

together_efn <- merge(ants_efn, plants_efn, by="abs")
colnames(together_efn) <- c("abs", "ants", "plants")

write.csv(ants_efn, file = "efn_ants_2.csv")
write.csv(plants_efn, file = "efn_plants_2.csv")

together_efn$abs <- NULL

final_table_efn <- table(together_efn)
final_table_efn <- (final_table_efn > 0) + 0
write.csv(final_table_efn, file = "efn_01_network.csv")
```


## DOMATIA
```{r}
dom_network <- read.csv("dom_network_tm_processed_2.csv", header=TRUE)
dom_network2 <- dom_network[dom_network$abs %in% dom_network$abs[duplicated(dom_network$abs)],]
dom_network2$X <- NULL

ants_dom <- dom_network2 %>% filter(name %in% ant_list$ants)
ants_dom <- remove.factors(ants_dom)
plants_dom <- dom_network2 %>% filter(name %in% angiosperms_list$plants)
plants_dom <- remove.factors(plants_dom)

write.csv(ants_dom, file = "domatia_ants_2.csv")
write.csv(plants_dom, file = "domatia_plants_2.csv")

together_dom <- merge(ants_dom, plants_dom, by="abs")
colnames(together_dom) <- c("abs", "ants", "plants")

together_dom$abs <- NULL

final_table_dom <- table(together_dom)
final_table_dom <- (final_table_dom > 0) + 0
write.csv(final_table_dom, file = "dom_01_network.csv")
```

